Peaks Over Threshold for Bursty Time Series

### 2018-04-18 article_springer/article_springer.Rmd
diagnostics section

The Theorem in the [previous Section](#sec:scaling) implies that for high 
$\ell$ we may 
approximate the distribution of $T(\ell)$ with an 
${\rm ML}(\beta, b(1/p))$ distribution.
Building on the POT (Peaks Over Threshold) method, we propose the 
following estimation procedure for the distribution of $T(\ell)$:

1. For varying thresholds $\ell \uparrow x_R$, extract datasets of
  exceedance times $\{T(\ell, i)\}_i$.

2. Fit a Mittag-Leffler distribution to each dataset, which
  results in the $\ell$-dependent estimates $\{\hat\beta(\ell)\}_\ell$ and 
  $\{\hat \sigma(\ell)\}_\ell$.

@@

The above approach, though theoretically sound, benefits from the following
practical adjustments:

@@

The estimates $\hat \beta$ and $\hat \sigma_0$ give an estimate of the 
distribution of exceedance times, dependent on the threshold $\ell$:
\begin{align*}
T(\ell) \sim {\rm ML}(\hat \beta, k^{-1/\hat \beta} \hat \sigma_0).
\end{align*}
For quick estimates of the Mittag-Leffler parameters the method of
log-transformed moments by @Cahoy2013 was used.

@@

The CTRM model is based on three main assumptions, which are repeated 
below. For each assumption, we suggest one means of checking if it holds:

I.i.d.:
: The pairs $(W_i, J_i)$ are i.i.d. An indication if this is true is given 
  by an auto-correlation plot for the logarithms (to ensure finite moments) 
  of the two time series.

Uncoupled:
: Each $W_i$ is independent of each $J_i$. We propose an empirical copula
  plot to check for any dependence.

Heavy tails:
: The waiting times are heavy-tailed with tail parameter 
  $\beta \in (0,1)$. A Hill plot can confirm this assumption.

The motivation behind a CTRM model is to fit extreme observations
and their inter-arrival times. 
Observations below a certain threshold may be interpreted as noise and 
discarded.
Since the signal rather than the noise needs to satisfy our modelling 
assumptions, we suggest that the above model assumptions are checked 
*after* the noisy small observations are discarded.

@@

```{r solar-flare-tail-scale, message=FALSE, fig.cap="\\label{fig:flares}Stability plots for the tail and scale parameter of the Mittag-Leffler distribution of the Solar Flare dataset."}
thin_flares_ctrm <- thin(ctrm = flares_ctrm, k = 150)
par(mfrow=c(1,2))
MLestimates(thin_flares_ctrm, tail = tail, scale = 2.5 * 1E7)
```

>>>>

```{r solar-flare-tail-scale, message=FALSE, fig.height=4.5, fig.width=8, fig.cap="\\label{fig:flares}Stability plots for the tail and scale parameter of the Mittag-Leffler distribution of the Solar Flare dataset. Dotted horizontal lines are at $\\beta = 0.85$ and $\\sigma_0 = 3 \\times 10^7$."}
thin_flares_ctrm <- flares %>% ctrm() %>% thin(k = 700)
par(mfrow=c(1,2))
MLestimates(thin_flares_ctrm, tail = 0.9, scale = 3 * 1E7)
```

The Theorem in Section \ref{sec:scaling} implies that for high 
threshold $\ell$ we may 
approximate the distribution of $T(\ell)$ with an 
${\rm ML}(\beta, b(1/p))$ distribution, 
where $b(\,)$ varies regularly at $\infty$ with parameter $1/\beta$.
Building on the POT (Peaks Over Threshold) method, we propose the 
following estimation procedure for the distribution of $T(\ell)$:

1. For a range of thresholds $\ell$ near the largest order statistics, 
  extract datasets of exceedance times $\{T(\ell, i)\}_i$.

2. For each choice of threshold $\ell$, 
  fit a Mittag-Leffler distribution to the resulting dataset
  $\{T(\ell, i)\}_i$. 
  This results in the estimates $\{\hat\beta(\ell)\}_\ell$ and 
  $\{\hat \sigma(\ell)\}_\ell$.

@@

The above approach, though theoretically sound, benefits from the following
practical adjustments (compare with Figure \ref{fig:flares}):

@@

The estimates $\hat \beta$ and $\hat \sigma_0$ give an estimate of the 
distribution of exceedance times, dependent on the threshold $\ell$:
\begin{align*}
T(\ell) \sim {\rm ML}(\hat \beta, k^{-1/\hat \beta} \hat \sigma_0).
\end{align*}
For quick estimates of the Mittag-Leffler parameters we have used 
the method of log-transformed moments by @Cahoy2013.

@@

```{r flare-diagnostics-1, fig.height=7, fig.cap="Diagnostic plots for the solar flare data: auto-correlation function.\\label{fig:flare-diagnostics-1}"}
flares_ctrm <- flares %>% ctrm() %>% thin(k = 200)
acf(flares_ctrm)
```

```{r flare-diagnostics-2, fig.height=7, fig.cap="Diagnostic plots for the solar flare data: empirical copula and QQ Plot. \\label{fig:flare-diagnostics-2}"}
par(mfrow = c(2,2))
empcopula(flares_ctrm)
mlqqplot(flares_ctrm, log = 'xy', tail = 0.8, main = "Mittag-Leffler QQ Plot")
```

The MRP model is based on three main assumptions, which are repeated 
below. For each assumption, we suggest one means of checking if it holds:

I.i.d.:
: After removing the "noise observations" below the smallest threshold 
  $\ell_0$, the pair sequence $(T(\ell_0, i), X(\ell_0,i))$ is i.i.d. 
  An indication if this is true is given 
  by an auto-correlation plot for the logarithms (to ensure finite moments) 
  of the two time series.

Uncoupled:
: Each $T(\ell, i)$ is independent of each $X(\ell, i)$. We propose an empirical copula
  plot to check for any dependence.

${\rm ML}(\beta, \sigma)$ distribution of $T(\ell, i)$:
: Apply a cutoff at the lowest threshold $\ell_0$, 
  extract the threshold crossing times, 
  and create a QQ Plot for the Mittag-Leffler distribution. 
  Use a log-Moment estimate of the tail parameter for the theoretical / population 
  quantiles of the plot.

Figures \ref{fig:flare-diagnostics-1} and \ref{fig:flare-diagnostics-2} 
show the diagnostic plots for a 
minimum threshold chosen at the 200th order statistic. 
There is some residual autocorrelation for the sequence of threshold 
exceedance times that is not accounted for by the MRP model. 
The fit with a Mittag-Leffler distribution ($\beta = 0.8$) is good, and 
there is no apparent dependence between threshold exceedance times and 
event magnitudes.

@@
